---
title: "IAS_COVID paper"
author: "Ihoghosa Iyamu"
date: "11/12/2020"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document:
    latex_engine : lualatex
    toc: yes
    toc_depth: 2
  html_notebook:
    toc: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MatchIt)
require(SASxport)
require(kableExtra)
require(knitr)
require(car)
require(DataExplorer)
require("tableone")
require(Publish)
require(mice)
require(jtools)
require(ggstance)
require(broom.mixed)
require(huxtable)
require(dplyr)
require(tidyverse)
require(cobalt)
require(survminer)
```

# Research Question 

This study aims to assess if social media as a main source of information on COVID-19 is associated with perceived efficacy of face masks 

## PICOT Table 

| PICOT | Definition |
|-|-|
| Population of interest | Adults aged 18 and above in 6 sub-saharan countries |
| Exposure of interest | Use of social media as main source of information on COVID-19 |
| Control group | Not using social media as a main source of information on COVID-19 |
| Outcomes of interest | Self-percieved risk of infection and efficacy of face masks | 
| Timeframe | 2020 |
|-|-|

# Data import and preparation
```{r data import}
require (foreign)
setwd("C:/Users/miyamu/OneDrive/Documents/Personal/COVID-19 papers")
PARC <- read.spss("PARC_survey.sav", use.value.label=TRUE, to.data.frame=TRUE)
##create analytic dataset 
analytic.data<-select(PARC,
                      Timestamp,
                      Sex,
                      Howwouldyoudescribeyourplaceofresidence, #residence
                      Country,
                      Confidence_current, #confidence in government response
                      Age_group,
                      Whatisthehighestlevelofeducationyouattained, #Level of education
                      WhatisyouremploymentstatusIfmorethanonepleaserespondusingthemain, #employment
                      HowdidyoufirstgettolearnaboutCOVID19,
                      TV,
                      Radio,
                      SocialMedia,
                      Friends,
                      Familyrelatives,
                      Onlineweb,
                      Employer,
                      Newspaper,
                      Others,
                      First_reaction_COVID,
                      Risk_Percept_COVID,
                      Facemask_efficacy)
```

```{r rename variables}
#rename the variables
names(analytic.data)<-c("ID",
                    "Sex", "Residence", "Country",
                    "Confidence.Response", "Age.category", "Level.of.education","Occupation",
                    "First.info", "TV", "Radio", "Social.media",
                    "Friends", "Family.relatives", "Online.Web", "Employer",
                    "Newspaper", "Others", "Attitude.COVID", "Risk.perception", "Facemask.efficacy")

#make factor variables 
var.names<-c("Sex", "Residence", "Level.of.education", "First.info","Occupation",
             "TV", "Radio", "Social.media", "Friends", "Family.relatives", 
             "Online.Web", "Employer", "Newspaper","Attitude.COVID", "Others")
analytic.data [var.names] <- lapply(analytic.data[var.names] , factor)
str(analytic.data)
```
## Recode factor levels labeled '0' and '1'
```{r}
analytic.data$TV<-ifelse(analytic.data$TV == 0, "No", "Yes")
table(analytic.data$TV, useNA = "always")

analytic.data$Radio<-ifelse(analytic.data$Radio == 0, "No", "Yes")
table(analytic.data$Radio, useNA = "always")

analytic.data$Social.media<-ifelse(analytic.data$Social.media == 0, "No", "Yes")
table(analytic.data$Social.media, useNA = "always")

analytic.data$Friends<-ifelse(analytic.data$Friends == 0, "No", "Yes")
table(analytic.data$Friends, useNA = "always")

analytic.data$Family.relatives<-ifelse(analytic.data$Family.relatives == 0, "No", "Yes")
table(analytic.data$Family.relatives, useNA = "always")

analytic.data$Online.Web<-ifelse(analytic.data$Online.Web == 0, "No", "Yes")
table(analytic.data$Online.Web, useNA = "always")

analytic.data$Employer<-ifelse(analytic.data$Employer == 0, "No", "Yes")
table(analytic.data$Employer, useNA = "always")

analytic.data$Newspaper<-ifelse(analytic.data$Newspaper == 0, "No", "Yes")
table(analytic.data$Newspaper, useNA = "always")

analytic.data$Infosource.others<-ifelse(analytic.data$Others == 0, "No", "Yes")
table(analytic.data$Infosource.others, useNA = "always")
```
## Recode other variable levels 
```{r}

#1. Level of education 
analytic.data$Level.of.education<-car::recode(analytic.data$Level.of.education, 
                                "c('Primary  ','Secondary')='Primary/Secondary'; 
                                 c('Tertiary ') = 'Tertiary'; else=NA")
table(analytic.data$Level.of.education, useNA = "always")

#2. Attitude to COVID 
analytic.data$Attitude.COVID<-car::recode(analytic.data$Attitude.COVID,
                                          "c('Worry')='Worry';
                                          c('Fear')='Fear';
                                          c('Calm')='Calm';
                                          c('Doubt')='Doubt';
                                          c('Disinterest','Others')='Others'")
table(analytic.data$Attitude.COVID, useNA = "always")

#3. Occupation
analytic.data$Occupation<-car::recode(analytic.data$Occupation,
                                       "c('Government employed                         ', 'Hospital-based worker                       ',
                                       'Non-governmental organization (NGO) employed', 'Private sector employed                     ',
                                       'Self employed                               ') = 'Employed';
                                       c('Retired                                     ', 'Unemployed                                  ') = 'Unemployed/retired'; 
                                      c('Student                                     ')='Student'")
```



```{r}
#Check missingness
require(DataExplorer)
plot_missing(analytic.data)
```

**Respondents who indicated "don't know" or "does not apply to my country" were excluded from the analysis**
```{r}
#complete case analysis
analytic.data1<-subset(analytic.data, Facemask.efficacy != "NA")
#exclude Tanzania
analytic.data2<-subset(analytic.data1, Country != "Tanzania")
plot_missing(analytic.data2)
```
## Drop unused levels
```{r}
analytic.data2$Country<-droplevels(analytic.data2$Country)
```

# Descriptive analysis 

```{r}
require(tableone)
Table1<-(CreateTableOne(vars = c("Facemask.efficacy"),
             data = analytic.data2, includeNA = TRUE, 
             test = FALSE, addOverall = TRUE))
PrintTable1<-print(Table1, showAllLevels = TRUE)
```


```{r pressure, echo=FALSE, results='hide'}
require(tableone)
Table2<-(CreateTableOne(vars = c("Facemask.efficacy","Sex", "Residence", "Country",
                    "Confidence.Response", "Age.category", "Level.of.education",
                    "TV", "Radio", "Occupation",
                    "Friends", "Family.relatives", "Online.Web", "Employer",
                    "Newspaper", "Infosource.others", "Risk.perception", "Attitude.COVID"),
             data = analytic.data2, strata = 'Social.media', includeNA = TRUE, 
             test = FALSE, addOverall = TRUE))
PrintTable2<-print(Table2, showAllLevels = TRUE)
```
```{r}
require(kableExtra)
kable(PrintTable2, booktabs = TRUE) %>%
kable_styling(latex_options = "scale_down")
```

# Bivariate analyis 
```{r}
form1<-as.formula(I(Facemask.efficacy=="Agree")~Social.media)
fit1 <- glm(form1, data = analytic.data2, family = binomial(logit))
publish(fit1)
```
```{r}
form2<-as.formula(I(Facemask.efficacy=="Agree")~Sex)
fit2 <- glm(form2, data = analytic.data2, family = binomial(logit))
publish(fit2)
```
```{r}
form3<-as.formula(I(Facemask.efficacy=="Agree")~Residence)
fit3 <- glm(form3, data = analytic.data2, family = binomial(logit))
publish(fit3)
```
```{r}
form4<-as.formula(I(Facemask.efficacy=="Agree")~Country)
fit4 <- glm(form4, data = analytic.data2, family = binomial(logit))
publish(fit4)
```
```{r}
form5<-as.formula(I(Facemask.efficacy=="Agree")~Confidence.Response)
fit5 <- glm(form5, data = analytic.data2, family = binomial(logit))
publish(fit5)
```
```{r}
form6<-as.formula(I(Facemask.efficacy=="Agree")~Age.category)
fit6 <- glm(form6, data = analytic.data2, family = binomial(logit))
publish(fit6)
```
```{r}
form7<-as.formula(I(Facemask.efficacy=="Agree")~Level.of.education)
fit7 <- glm(form7, data = analytic.data2, family = binomial(logit))
publish(fit7)
```
```{r}
form8<-as.formula(I(Facemask.efficacy=="Agree")~Occupation)
fit8 <- glm(form8, data = analytic.data2, family = binomial(logit))
publish(fit8)
```

```{r}
form9<-as.formula(I(Facemask.efficacy=="Agree")~TV)
fit9 <- glm(form9, data = analytic.data2, family = binomial(logit))
publish(fit9)
```
```{r}
form10<-as.formula(I(Facemask.efficacy=="Agree")~Radio)
fit10 <- glm(form10, data = analytic.data2, family = binomial(logit))
publish(fit10)
```
```{r}
form11<-as.formula(I(Facemask.efficacy=="Agree")~Friends)
fit11 <- glm(form11, data = analytic.data2, family = binomial(logit))
publish(fit11)
```

```{r}
form12<-as.formula(I(Facemask.efficacy=="Agree")~Family.relatives)
fit12 <- glm(form12, data = analytic.data2, family = binomial(logit))
publish(fit12)
```
```{r}
form13<-as.formula(I(Facemask.efficacy=="Agree")~Online.Web)
fit13 <- glm(form13, data = analytic.data2, family = binomial(logit))
publish(fit13)
```
```{r}
form14<-as.formula(I(Facemask.efficacy=="Agree")~Employer)
fit14 <- glm(form14, data = analytic.data2, family = binomial(logit))
publish(fit14)
```
```{r}
form15<-as.formula(I(Facemask.efficacy=="Agree")~Newspaper)
fit15 <- glm(form15, data = analytic.data2, family = binomial(logit))
publish(fit15)
```
```{r}
form16<-as.formula(I(Facemask.efficacy=="Agree")~Risk.perception)
fit16 <- glm(form16, data = analytic.data2, family = binomial(logit))
publish(fit16)
```
```{r}
form17<-as.formula(I(Facemask.efficacy=="Agree")~Attitude.COVID)
fit17 <- glm(form17, data = analytic.data2, family = binomial(logit))
publish(fit17)
```
# Multivariate analyses
## Step-wise AIC model
A step-wise AIC based model (with backward elimination) was fit after including a set of covariates identified in the literature to be associated with percieved efficacy of COVID-19 measures including mandatory mask wearing.
**Citations:** [Prasetyo et. al, 2020](https://doi.org/10.1016/j.ijid.2020.07.074) & [Howard, 2020](https://doi.org/10.1016/j.paid.2020.110417)
```{r}
#stepwise but with a scope
formula1 <- as.formula(I(Facemask.efficacy=="Agree")~Social.media+Sex+Residence+Country+
                         Confidence.Response+Age.category+Level.of.education+Occupation+
                         TV+Radio+Friends+Family.relatives+Online.Web+Employer+Newspaper+
                         Risk.perception+Attitude.COVID)
                    
scope <- list(upper =~Social.media+Sex+Residence+Country+
                         Confidence.Response+Age.category+Level.of.education+Occupation+
                         TV+Radio+Friends+Family.relatives+Online.Web+Employer+Newspaper+
                         Risk.perception+Attitude.COVID,
              lower = ~Social.media
                     +Risk.perception+Attitude.COVID+Age.category+Sex+Level.of.education+TV)
fita <- glm(formula1, data = analytic.data2, 
              family = binomial(logit))
fitstep <- step(fita, scope = scope, trace = FALSE,
                k = 2, direction = "backward")
publish(fitstep)
```
## Assessing possible interactions 
To the best of my knowledge, no evidence of model interactions. However, gender and age may be effect modifiers. Best to check
*Recheck step with another test after review*
```{r}
fitb <- update(fitstep, .~. + interaction(Social.media,Sex))
anova(fitstep, fitb, test = "LRT") # do not keep interaction
```
```{r}
fitc <- update(fitstep, .~. + interaction(Social.media,Age.category))
anova(fitstep, fitc, test = "LRT") # do not keep interaction
```
## Final Model
```{r}
publish(fitstep)
vif(fitstep)
```
## Model diagnostics 
Not necessarily important since it is not a predictive model
```{r}
require(pROC)
pred.y <- predict(fitstep, type = "response")
rocobj <- roc(analytic.data2$Facemask.efficacy, pred.y)
rocobj
auc(rocobj)
```
```{r, Hosmer-lemeshow test}
require(ResourceSelection)
hl<-hoslem.test(fitstep$y, fitted(fitstep))
hl
```
There is no evidence of a lack of fit in this model

#Propensity score analysis
##issue unique ID per row
```{r}
analytic.data2$ID1 <- 1:nrow(analytic.data2)
```

## Prepare the data 
```{r, results='hide'}
str(analytic.data2)
dim(analytic.data2)
var.names<-c("TV","Radio","Social.media","Friends", "Family.relatives", 
             "Online.Web", "Employer", "Newspaper", "Infosource.others")
analytic.data2[var.names] <- lapply(analytic.data2[var.names] , factor)
str(analytic.data2)
# ...
```
##Format variables 
```{r, results='hide'}

# Outcome recode
analytic.data2$Facemask.efficacy <- ifelse(analytic.data2$Facemask.efficacy == "Agree", 1, 0)

# Exposure recode
analytic.data2$Social.media <- ifelse(analytic.data2$Social.media == "Yes", 1, 0) 

dim(analytic.data2)

```


## Set seed
```{r}
# set a seed
set.seed(346)
```

# (optional) Propensity score matching analysis by Austin P. C., Jembere N. and Chiu M. (2018)

## Step 1

```{r}
##
# response = exposure variable
# independent variables = baseline covariates
ps.formula <- as.formula(Social.media~Sex+Residence+Country+
                         Confidence.Response+Age.category+Level.of.education+Occupation+
                         TV+Radio+Friends+Family.relatives+Online.Web+Employer+Newspaper+
                         +Others+Risk.perception+Attitude.COVID+First.info+Infosource.others)

ps.fit <- glm(ps.formula, data = analytic.data2, family=binomial(logit))
analytic.data2$PS <- fitted(ps.fit)
summary(analytic.data2$PS)
```
## Step 2


```{r}
##
require(Matching)
match.obj2 <- Match(Y=analytic.data2$Facemask.efficacy, 
                    Tr=analytic.data2$Social.media, 
                    X=analytic.data2$PS, 
                    M=1, 
                    estimand = "ATT",
                    replace=FALSE, 
                    caliper = 0.2)
summary(match.obj2)
matched.data2 <- analytic.data2[c(match.obj2$index.treated, 
                                 match.obj2$index.control),]
dim(matched.data2)
```

```{r, cache=TRUE}
require(MatchIt)
set.seed(346)
match.obj <- matchit(ps.formula, data = analytic.data2,
                     distance = analytic.data2$PS, 
                     method = "nearest", 
                     replace=FALSE,
                     caliper = .2, 
                     ratio = 1)
analytic.data2$PS <- match.obj$distance
summary(match.obj$distance)
plot(match.obj, type = "jitter")
plot(match.obj, type = "hist")
tapply(analytic.data2$PS, analytic.data2$Social.media, summary)
# check how many matched
match.obj
# extract matched data
matched.data2 <- match.data(match.obj)
dim(matched.data2)
```
## Step 3

```{r}
##
baselinevars <- c("Sex", "Residence", "Country",
                    "Confidence.Response", "Age.category", "Level.of.education","Occupation",
                    "TV", "Radio", "Friends", "Family.relatives", "Online.Web", "Employer",
                    "Newspaper", "Others", "Attitude.COVID", "Risk.perception", "Facemask.efficacy")
tab1m <- CreateTableOne(strata = "Social.media", 
                           vars = baselinevars,
                           data = analytic.data2, test = FALSE)
print(tab1m, smd = TRUE)
```

```{r}
tab2m <- CreateTableOne(strata = "Social.media", 
                           vars = baselinevars,
                           data = matched.data2, test = FALSE)
print(tab2m, smd = TRUE)
```

## Step 4

```{r}
##
analytic.data2$matched <- 0
length(analytic.data2$ID) # full data
length(matched.data2$ID) # matched data
length(analytic.data2$ID[analytic.data2$ID %in% matched.data2$ID])
analytic.data2$matched[analytic.data2$ID %in% matched.data2$ID] <- 1
table(analytic.data2$matched)

```

```{r, warning=FALSE}
out.formula <- as.formula(Facemask.efficacy ~ Social.media)
ps.fit2 <- glm(out.formula, data = matched.data2, family = binomial(logit))
publish(ps.fit2)
```



